Structural Divergence - Protein dynamics

### 2017-06-25 analysis-MSF-SD-RSD.Rmd
23/6/17

---
title: "Dynamical Divergence Analysis"
author: "Maria Laura Marcos, Julian Echave"
output: html_document
---

# Introduction
We study the evolutionary divergence of proteins using the LF-ENM and comparing simulated mutants with experimental data.

```{r setup, include = FALSE}
# set chunk options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```
```{r load-packages}
# load packages
library(ggplot2)
library(reshape2)
library(plyr)
library(quantreg)
library(bio3d)

@@

## profiles to compare
### square.dif.MSF
square.dif.MSF.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.square.dif.MSF.csv", sep = ""))
square.dif.MSF.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.square.dif.MSF.csv", sep = ""))
square.dif.MSF.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.square.dif.MSF.csv", sep = ""))

### dri2
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
dri2.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))

```
```{r set-reference-protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```
We studied the `r family` family. 
We used the ENM `r enm` with R0 = `r R0` to generate mutants.
We used as reference for the ENM the protein with pdb code `r reference`. 
We simulated `r n.mut.p` mutants of this reference for each protein of the family.
The `r family` family is of the type `r type` and consists of the following proteins:

@@

# read reference pdb and calculate resno, AA, and CN
reference.pdb.fname = file.path(data.dir,paste(reference, ".pdb", sep = ""))
ref.pdb = read.pdb(reference.pdb.fname)
inds = atom.select(ref.pdb, elety = "CA")
ref.pdb.ca = ref.pdb$atom[inds$atom,]
cmap.ca = cmap(ref.pdb$xyz[inds$xyz], dcut = R0, scut = 0, mask.lower = F)
r.p.ref = ref.pdb$xyz[inds$xyz]
CN = rowSums(cmap.ca, na.rm = T) - 1 # diag(cmap.ca) returned by cmap is 1!
n.sites = length(CN)
AA = ref.pdb.ca$resid
resno = ref.pdb.ca$resno
pdb.info = data.frame(resno, AA, CN)
names(pdb.info) = c("resno", "AA", paste("CN", as.character(R0), sep = ""))

@@

```{r read-site-data}
# read MSF data
square.dif.MSF.exp = read.csv(square.dif.MSF.exp.fname, header = T)
square.dif.MSF.mut = read.csv(square.dif.MSF.mut.fname, header = T)
square.dif.MSF.medium = read.csv(square.dif.MSF.medium.fname, header = T)

# read dri2 data
dri2.exp = read.csv(dri2.exp.fname, header = T)
dri2.mut = read.csv(dri2.mut.fname, header = T)
dri2.medium = read.csv(dri2.medium.fname, header = T)
```

```{r MSF-p.ref}
# calculate MSF (Mean Square Fluctuation)

## calculate ENM
ENMK.p.ref = CalculateENMKeff(matrix(r.p.ref, nrow = 3), 
                              seq(1:n.sites), 
                              c(), 
                              R0, 
                              tolerance, 
                              K.analysis)
## get cov matrix
cov.p.ref = ENMK.p.ref$cov

@@

```{r prepare-site-data-MSF}
# prepare the data
## exp
tsquare.dif.MSF.exp = as.data.frame(t(as.matrix(square.dif.MSF.exp)))
tsquare.dif.MSF.exp = cbind(site.info, tsquare.dif.MSF.exp)
siteComparison.exp = melt(tsquare.dif.MSF.exp, id.vars = names(site.info), variable.name = "protein", value.name = "square.dif.MSF")

## mut
tsquare.dif.MSF.mut = as.data.frame(t(as.matrix(square.dif.MSF.mut)))
tsquare.dif.MSF.mut = cbind(site.info, tsquare.dif.MSF.mut)
siteComparison.mut = melt(tsquare.dif.MSF.mut, id.vars = names(site.info), variable.name = "protein", value.name = "square.dif.MSF")

## medium
tsquare.dif.MSF.medium = as.data.frame(t(as.matrix(square.dif.MSF.medium)))
tsquare.dif.MSF.medium = cbind(site.info, tsquare.dif.MSF.medium)
siteComparison.medium = melt(tsquare.dif.MSF.medium, id.vars = names(site.info), variable.name = "protein", value.name = "square.dif.MSF")

# bluild the data.frame
siteComparison.all.datasets = rbind(data.frame("dataset" = "exp", siteComparison.exp),
                                    data.frame("dataset" = "mut", siteComparison.mut),
                                    data.frame("dataset" = "medium", siteComparison.medium))

siteComparison.all.datasets = ddply(siteComparison.all.datasets, c("dataset", "protein"), mutate,
          "r.SDF" = rank(square.dif.MSF, na.last = "keep")) 
```

@@

siteComparison.all.datasets.dri2 = ddply(siteComparison.all.datasets.dri2, c("dataset", "protein"), mutate,
          "RSD" = sqrt(dri2), 
        "z.RSD" = vscale(RSD))
```

### Dynamical divergence along sites

In this section we consider dynamical divergence along Cartessian coordinates. square.dif.MSF is the square diference between MSF of the proteins relative to the reference protein. We ranked profiles because we found it is the best way to deal with large differences in MSF of divergent sites. We also tried z-normalizing but the results were not good.

```{r mean-se-r.SDF-vs-site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}

# comparison of r.SDF profiles with functional sites
## prepare dat
dat = siteComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("medium" = "selection"))
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))

## dat with means and standard errors
d = ddply(dat,c("dataset", "site", "AA", "shell"), function(x) {
                mean.r.SDF = mean(x$r.SDF, na.rm = T)
                n.data = sum(!is.na(x$r.SDF))
                se.r.SDF = sd(x$r.SDF, na.rm = T) / sqrt(n.data)
                data.frame(mean.r.SDF, n.data, se.r.SDF)
})

## find the active site and its contacts
active.sites = d$site[d$shell == 0]
contacts.active.sites = d$site[d$shell == 1]

## prepare plots with mean profiles and deviations
p = ggplot(d, aes(x = site, y = mean.r.SDF, col = dataset, fill = dataset)) +
    geom_line() +
    geom_ribbon(aes(ymin = mean.r.SDF - f95 * se.r.SDF, ymax = mean.r.SDF + se.r.SDF), alpha = .4, show.legend = F) +
    geom_vline(xintercept = active.sites, show.legend = F) +
    geom_vline(xintercept = contacts.active.sites, col = "grey", show.legend = F) +
    labs(x = "site", y = "<ranked square dif MSFi>") +
    scale_colour_manual(name = NULL, values = myColors) + 
    scale_fill_manual(name = NULL, values = myColors)

## plot separate and comparative plots
p.facet = p + facet_grid(dataset~., switch = "y") + theme(legend.position = "bottom")

p.facet %+% subset(d, dataset == "exp") 
p.facet %+% subset(d, dataset == "no-selection")
p.facet %+% subset(d, dataset == "selection") 
p.facet

## plot superimposed plots
p + theme(legend.position = "bottom")
```

```{r comparison-of-means}

@@

mean.r.square.dif.MSF.exp = d.exp$mean.r.SDF
mean.r.square.dif.MSF.mut = d.mut$mean.r.SDF
mean.r.square.dif.MSF.medium = d.medium$mean.r.SDF

# calculate correlations
r.exp.mut = cor(mean.r.square.dif.MSF.exp, mean.r.square.dif.MSF.mut)
r.exp.medium = cor(mean.r.square.dif.MSF.exp, mean.r.square.dif.MSF.medium)
r.mut.medium = cor(mean.r.square.dif.MSF.mut, mean.r.square.dif.MSF.medium)

cor.df = data.frame("comparison" = c("exp vs no-selection", "exp vs selection", "no-selection vs selection"),
                    "R" = c(r.exp.mut, r.exp.medium, r.mut.medium))

knitr::kable(cor.df, digits = 2)

# plot comparative profiles 
par(pty = "s")

plot(x = mean.r.square.dif.MSF.exp, 
     y = mean.r.square.dif.MSF.mut, 
     xlab = "<r.square.dif.MSF> experimental", 
     ylab = "<r.square.dif.MSF> no-selection")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.exp, 
     y = mean.r.square.dif.MSF.medium, 
     xlab = "<r.square.dif.MSF> experimental", 
     ylab = "<r.square.dif.MSF> selection")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.mut, 
     y = mean.r.square.dif.MSF.medium, 
     xlab = "<r.square.dif.MSF> no-selection", 
     ylab = "<r.square.dif.MSF> selection")
abline(0, 1)
```

```{r comparison-with-MSF.p.ref}

# rank MSF.p.ref
r.MSF.p.ref = rank(MSF.p.ref)

# calculate correlations
r.exp = cor(mean.r.square.dif.MSF.exp, r.MSF.p.ref)
r.mut = cor(mean.r.square.dif.MSF.mut, r.MSF.p.ref)
r.medium = cor(mean.r.square.dif.MSF.medium, r.MSF.p.ref)

cor.df = data.frame("comparison" = c("<r.square.dif.MSF> experimental vs r.MSF.p.ref", "<r.square.dif.MSF> no-selection vs r.MSF.p.ref", "<r.square.dif.MSF> selection vs r.MSF.p.ref"),
                    "R" = c(r.exp, r.mut, r.medium))

knitr::kable(cor.df, digits = 2)

# plot comparative profiles 
par(pty = "s")

plot(x = mean.r.square.dif.MSF.exp, 
     y = r.MSF.p.ref, 
     xlab = "<r.square.dif.MSF> experimental", 
     ylab = "r.MSF.p.ref")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.mut, 
     y = r.MSF.p.ref, 
     xlab = "<r.square.dif.MSF> no-selection", 
     ylab = "r.MSF.p.ref")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.medium, 
     y = r.MSF.p.ref, 
     xlab = "<r.square.dif.MSF> selection", 
     ylab = "r.MSF.p.ref")
abline(0, 1)
```
```{r mean-z.RSD-profiles}

# prepare dat
dat = siteComparison.all.datasets.dri2
dat$dataset = revalue(dat$dataset, c("medium" = "selection"))
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))

# dat with means and standard errors
d = ddply(dat,c("dataset", "site", "AA", "shell"), function(x) {
                mean.z.RSD = mean(x$z.RSD, na.rm = T)
                n.data = sum(!is.na(x$z.RSD))
                se.z.RSD = sd(x$z.RSD, na.rm = T) / sqrt(n.data)
                data.frame(mean.z.RSD, n.data, se.z.RSD)
})

mean.r.dri2.exp = d.exp$mean.z.RSD
mean.r.dri2.mut = d.mut$mean.z.RSD
mean.r.dri2.medium = d.medium$mean.z.RSD
```

### Comparison structure - dynamics

In this section we compare dri2 profiles (structural variability) with square dif MSF profiles (dynamical variability). To do this, we ranked each profile.

```{r structure-vs-dynamics}

# rank profiles
mean.r.dri2.exp = rank(mean.r.dri2.exp)
mean.r.dri2.mut = rank(mean.r.dri2.mut)
mean.r.dri2.medium = rank(mean.r.dri2.medium)

mean.r.square.dif.MSF.exp = rank(mean.r.square.dif.MSF.exp)
mean.r.square.dif.MSF.mut = rank(mean.r.square.dif.MSF.mut)
mean.r.square.dif.MSF.medium = rank(mean.r.square.dif.MSF.medium)

# plot comparative profiles
par(pty = "s")

plot(x = mean.r.square.dif.MSF.exp, 
     y = mean.r.dri2.exp, 
     xlab = "r<r.square.dif.MSF> experimental", 
     ylab = "r<dri2> experimental")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.mut, 
     y = mean.r.dri2.mut, 
     xlab = "r<r.square.dif.MSF> no-selection", 
     ylab = "r<dri2> no-selection")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.medium, 
     y = mean.r.dri2.medium, 
     xlab = "r<r.square.dif.MSF> selection", 
     ylab = "r<dri2> selection")
abline(0, 1)

# calculate correlations
r.exp = cor(mean.r.square.dif.MSF.exp, mean.r.dri2.exp)
r.medium = cor(mean.r.square.dif.MSF.medium, mean.r.dri2.medium)
r.mut = cor(mean.r.square.dif.MSF.mut, mean.r.dri2.mut)

cor.df = data.frame("comparison structure - dynamics" = c("exp", "selection", "no-selection"),
                    "R" = c(r.exp, r.medium, r.mut))
knitr::kable(cor.df, digits = 2)
```

>>>>

---
title: "Dynamical Divergence Analysis"
author: "Maria Laura Marcos, Julian Echave"
output: html_document
---
# Introduction
In this report we compare structural divergence profiles in cartesian coordinates (RMSD and MSD) with the Mean Square Fluctuation (MSF) of the sites.

```{r setup, include = FALSE}
# set chunk options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load-packages}
# load packages
library(ggplot2)
library(reshape2)
library(plyr)
library(quantreg)
library(bio3d)

@@

## profiles to compare
### dri2
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
dri2.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
```

```{r set-reference-protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```
We studied the `r family` family. 
We used the ENM `r enm` with R0 = `r R0` to generate mutants.
We used as reference for the ENM the protein with pdb code `r reference`. 
We simulated `r n.mut.p` mutants of this reference for each protein of the family.
The `r family` family is of the type `r type` and consists of the following proteins:

@@

# read reference pdb and calculate resno, AA, and CN
reference.pdb.fname = file.path(data.dir,paste(reference, ".pdb", sep = ""))
ref.pdb = read.pdb(reference.pdb.fname)
inds = atom.select(ref.pdb, elety = "CA")
ref.pdb.ca = ref.pdb$atom[inds$atom,]
cmap.ca = cmap(ref.pdb$xyz[inds$xyz], dcut = R0, scut = 0, mask.lower = F)
r.p.ref = ref.pdb$xyz[inds$xyz]
CN = rowSums(cmap.ca, na.rm = T) - 1 # diag(cmap.ca) returned by cmap is 1!
n.sites = length(CN)
AA = ref.pdb.ca$resid
resno = ref.pdb.ca$resno

pdb.info = data.frame(resno, AA, CN)
names(pdb.info) = c("resno", "AA", paste("CN", as.character(R0), sep = ""))

@@

```{r read-site-data}
# read dri2 data
dri2.exp = read.csv(dri2.exp.fname, header = T)
dri2.mut = read.csv(dri2.mut.fname, header = T)
dri2.medium = read.csv(dri2.medium.fname, header = T)
```

```{r calculate-MSF-p.ref}
# calculate MSF of p.ref (Mean Square Fluctuation)
## calculate ENM
ENMK.p.ref = CalculateENMKeff(matrix(r.p.ref, nrow = 3), 
                              seq(1:n.sites), 
                              c(), 
                              R0, 
                              tolerance, 
                              K.analysis)
## get cov matrix
cov.p.ref = ENMK.p.ref$cov

@@

@@

siteComparison.all.datasets.dri2 = ddply(siteComparison.all.datasets.dri2, c("dataset", "protein"), mutate,
          "RSD" = sqrt(dri2), 
           "SD" = dri2, 
        "z.RSD" = vscale(RSD),
         "z.SD" = vscale(SD))
```

```{r mean-profiles-comparisons}
# prepare dat
dat = siteComparison.all.datasets.dri2
dat$dataset = revalue(dat$dataset, c("medium" = "selection"))
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))

# dat with means and standard errors
d = ddply(dat,c("dataset", "site", "AA", "shell"), function(x) {
                mean.z.RSD = mean(x$z.RSD, na.rm = T)
                mean.RSD = mean(x$RSD, na.rm = T)
                mean.z.SD = mean(x$z.SD, na.rm = T)
                mean.SD = mean(x$SD, na.rm = T)
                n.data = sum(!is.na(x$z.RSD))
                se.z.RSD = sd(x$z.RSD, na.rm = T) / sqrt(n.data)
                data.frame(mean.z.RSD,mean.RSD, mean.z.SD, mean.SD, n.data, se.z.RSD)
})

@@

# MSF.p.ref vs z.RSD and z.SD
MSF.p.ref = as.vector(MSF.p.ref)

# z.RSD
## correlations
r.exp = cor(d.exp$mean.RSD, MSF.p.ref)
r.medium = cor(d.mut$mean.RSD, MSF.p.ref)
r.mut = cor(d.medium$mean.RSD, MSF.p.ref)

cor.df = data.frame("MSF vs RSD" = c("exp", "selection", "no-selection"),
                             "R" = c(r.exp, r.medium, r.mut))
knitr::kable(cor.df, digits = 2)

## plots
par(pty = "s")

plot(y = d.exp$mean.RSD, x = MSF.p.ref, ylab = "RSD experimental")
lm = lm(d.exp$mean.RSD ~ MSF.p.ref)
abline(lm)
legend("topleft", paste ("r = ", round(r.exp, digits = 2)), bty = "n")

plot(y = d.mut$mean.RSD, x = MSF.p.ref, ylab = "RSD no-selection")
lm = lm(d.mut$mean.RSD ~ MSF.p.ref)
abline(lm)
legend("topleft", paste ("r = ", round(r.mut, digits = 2)), bty = "n")

plot(y = d.medium$mean.RSD, x = MSF.p.ref, ylab = "RSD selection")
lm = lm(d.medium$mean.RSD ~ MSF.p.ref)
abline(lm)
legend("topleft", paste ("r = ", round(r.medium, digits = 2)), bty = "n")

# SD
## correlations
r.exp = cor(d.exp$mean.SD, MSF.p.ref)
r.medium = cor(d.mut$mean.SD, MSF.p.ref)
r.mut = cor(d.medium$mean.SD, MSF.p.ref)

cor.df = data.frame("MSF vs SD" = c("exp", "selection", "no-selection"),
                            "R" = c(r.exp, r.medium, r.mut))
knitr::kable(cor.df, digits = 2)

## plots
par(pty = "s")

plot(y = d.exp$mean.SD, x = MSF.p.ref, ylab = "SD experimental")
lm = lm(d.exp$mean.SD ~ MSF.p.ref)
abline(lm)
legend("topleft", paste ("r = ", round(r.exp, digits = 2)), bty = "n")

plot(y = d.mut$mean.SD, x = MSF.p.ref, ylab = "SD no-selection")
lm = lm(d.mut$mean.SD ~ MSF.p.ref)
abline(lm)
legend("topleft", paste ("r = ", round(r.mut, digits = 2)), bty = "n")

plot(y = d.medium$mean.SD, x = MSF.p.ref, ylab = "SD selection")
lm = lm(d.medium$mean.SD ~ MSF.p.ref)
abline(lm)
legend("topleft", paste ("r = ", round(r.medium, digits = 2)), bty = "n")
```

### 2017-06-25 analysis-Pn-energy.Rmd
23/6/17

---
title: "Structural Divergence Analysis"
author: "Maria Laura Marcos, Julian Echave"
output: html_document
---

# Introduction
We study the evolutionary divergence of proteins using the LF-ENM and comparing simulated mutants with experimental data.

@@

```{r set_file_names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))
site.info.fname = file.path(data.dir, paste(p.ref, "_m.da.ca.csv", sep = ""))
min.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_min.da.CM.ca.csv", sep = ""))

@@

### energy
energy.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.va.csv", sep = ""))
energy.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.va.csv", sep = ""))
energy.mut.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.va.csv", sep = ""))

@@

Pn.medium = read.csv(Pn.medium.fname, header = T)
rownames(Pn.medium) = paste("medium", rownames(Pn.medium), sep = "")
energy.medium = read.csv(energy.medium.fname, header = T)
rownames(energy.medium) = paste("medium", rownames(energy.medium), sep = "")

### Structural divergence along normal modes

In this section we consider structural divergence using normal mode coordinates. Pn is the square divergence along a normal mode relative to the mean.

@@

modeComparison.exp = merge(energy.exp.long,Pn.exp.long,by=c("mode", "protein"))
# renormalize Pn
modeComparison.exp = ddply(modeComparison.exp, "protein", mutate, "Pn" = Pn / mean(Pn, na.rm = T))

# modeComparison.exp has NA's because for proteins with gaps when using Keff there're less modes
#print("Warning: modeComparison.exp <- na.omit(modeComparison.exp)")
modeComparison.exp <- na.omit(modeComparison.exp)
modeComparison.medium <- na.omit(modeComparison.medium)
modeComparison.mut <- na.omit(modeComparison.mut)
```

```{r add-quantile-regression-data}

## Quantile regression

# fit median and quantiles to simulated data and add them to data.frames
fit.medium.rq.5 <- rq(Pn ~ sn2, tau = .5, data = modeComparison.medium, method = "fn")
fit.medium.rq.05 <- rq(Pn ~ sn2, tau = .05, data = modeComparison.medium, method = "fn")
fit.medium.rq.95 <- rq(Pn ~ sn2, tau = .95, data = modeComparison.medium, method = "fn")
modeComparison.medium$rq.5 = fit.medium.rq.5$fitted.values
modeComparison.medium$rq.05 = fit.medium.rq.05$fitted.values
modeComparison.medium$rq.95 = fit.medium.rq.95$fitted.values

# fit median and quantiles to simulated data and add them to data.frames
fit.mut.rq.5 <- rq(Pn ~ sn2, tau = .5, data = modeComparison.mut, method = "fn")
fit.mut.rq.05 <- rq(Pn ~ sn2, tau = .05, data = modeComparison.mut, method = "fn")
fit.mut.rq.95 <- rq(Pn ~ sn2, tau = .95, data = modeComparison.mut, method = "fn")
modeComparison.mut$rq.5 = fit.mut.rq.5$fitted.values
modeComparison.mut$rq.05 = fit.mut.rq.05$fitted.values
modeComparison.mut$rq.95 = fit.mut.rq.95$fitted.values

# fit median and quantiles to experimental data
fit.exp.rq.5 <- rq(Pn ~ sn2, tau = .5, data = modeComparison.exp, method = "fn")
fit.exp.rq.05 <- rq(Pn ~ sn2, tau = .05, data = modeComparison.exp, method = "fn")
fit.exp.rq.95 <- rq(Pn ~ sn2, tau = .95, data = modeComparison.exp, method = "fn")
modeComparison.exp$rq.5 = fit.exp.rq.5$fitted.values
modeComparison.exp$rq.05 = fit.exp.rq.05$fitted.values
modeComparison.exp$rq.95 = fit.exp.rq.95$fitted.values

# put simulated and experimental data together

modeComparison.all.datasets = rbind(data.frame("dataset" = "exp", modeComparison.exp),
                                    data.frame("dataset" = "mut", modeComparison.mut),
                                    data.frame("dataset" = "medium", modeComparison.medium))

modeComparison.all.datasets$dataset = factor(modeComparison.all.datasets$dataset, levels = c("exp", "mut", "medium"))
```
### Pn vs. mode energy

Note: lines are the median, 0.05 and 0.95 quantiles fit using rq(Pn ~ sn2). sn2 = 1/energy is the expected MSD of along mode n. The mean value of Pn should be a linear function of sn2 (Pn is normalized, sn2 is not).

Here we show the square deviation along each normal mode (normalized so that = 1) vs. the normal-mode's energy. For the empirical data, we use the actual energy obtained from Keff for the given comparison. It is this energy the one that should be used, since it's inverse is the expected value of Pn. Lines are the .5 (median), .05, and .95 quantiles obtained using rq()

Upper panel: simulated vs empirical data, shown separately.
Lower panel: simulated vs empirical overlaped and zoomed to better see the pattern.

```{r Pn-vs-energy}

# Plot Pn vs energy
dat = modeComparison.all.datasets

# p is just the median and quantiles
p.points.rq = ggplot(dat, aes(x = energy, y = Pn, col = dataset)) + 
    geom_point(alpha = .4, size = .2) +
    geom_line(aes(y = rq.5)) +
    geom_line(aes(y = rq.05)) +
    geom_line(aes(y = rq.95)) 
p = p.points.rq +  
    labs(x = "mode energy", y = "structural divergence") +
    coord_cartesian(xlim = c(0,3)) +
    scale_colour_manual(name = NULL, values = myColors) + 
    scale_fill_manual(name=NULL, values = myColors)

p.facet = p + facet_grid(dataset~., switch = "y") + theme(legend.position = "none")
p.facet %+% subset(dat, dataset == "exp")
p.facet %+% subset(dat, dataset == "mut")
p.facet %+% subset(dat, dataset == "medium")
p.facet
```

### median(Pn) vs mode energy

@@

dat = ddply(dat,c("dataset", "mode"), function(x) {
    energy.mean = mean(x$energy)
    Pn.median = median(x$Pn)
    Pn.median.ci = sort(x$Pn)[qbinom(c(.025, .975), length(x$Pn), 0.5)]
    Pn.median.min = Pn.median.ci[1]
    Pn.median.max = Pn.median.ci[2]
    rq.5 = mean(x$rq.5)
    data.frame(energy.mean, Pn.median.min, Pn.median.max, Pn.median, rq.5)
})

p = ggplot(dat,aes(x=energy.mean,y=Pn.median,col=dataset,fill=dataset)) +
    geom_pointrange(aes(ymin=Pn.median.min,ymax = Pn.median.max),size=.2,alpha=.4) +
    geom_line(aes(y=rq.5))  +
    labs(x="mode energy",y="structural divergence (median Pn) ") +
    scale_colour_manual(name = NULL,values = myColors) + 
    scale_fill_manual(name=NULL,values=myColors)+
    coord_cartesian(xlim=c(0,3))

@@

```{r correlation-mode-profiles}
#dat$dataset = revalue(dat$dataset,c("medium" = "mut+sel"))
dat = modeComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("medium" = "selection"))
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))

# calculate dataset of medians
dat = ddply(dat, c("dataset", "mode"), function(x) {
    energy.mean = mean(x$energy)
    Pn.median = median(x$Pn)
    data.frame(energy.mean, Pn.median)
})

dat = subset(dat, mode < 101)
dat.exp = subset(dat, dataset == "exp")
dat.mut = subset(dat,dataset == "no-selection")
dat.medium = subset(dat,dataset == "selection")

r.exp.mut = cor(dat.exp$Pn.median, dat.mut$Pn.median)
r.exp.medium = cor(dat.exp$Pn.median, dat.medium$Pn.median)
r.mut.medium = cor(dat.mut$Pn.median, dat.medium$Pn.median)
cor.df = data.frame("comparison" = c("exp vs mut", "exp vs medium", "mut vs medium"),
                    "R" = c(r.exp.mut, r.exp.medium, r.mut.medium))
knitr::kable(cor.df, digits = 2)

# plot exp vs simulated profiles 
par(pty = "s")

plot(x = dat.exp$Pn.median, 
     y = dat.mut$Pn.median, 
     xlab = "exp.Pn.median", 
     ylab = "mut.Pn.median", 
     xlim = c(min(c(dat.exp$Pn.median, dat.mut$Pn.median)), max(c(dat.exp$Pn.median, dat.mut$Pn.median))),
     ylim = c(min(c(dat.exp$Pn.median, dat.mut$Pn.median)), max(c(dat.exp$Pn.median, dat.mut$Pn.median))))
abline(0, 1)

plot(x = dat.exp$Pn.median, 
     y = dat.medium$Pn.median, 
     xlab = "exp.Pn.median", 
     ylab = "medium.Pn.median", 
     xlim = c(min(c(dat.exp$Pn.median, dat.medium$Pn.median)), max(c(dat.exp$Pn.median, dat.medium$Pn.median))),
     ylim = c(min(c(dat.exp$Pn.median, dat.medium$Pn.median)), max(c(dat.exp$Pn.median, dat.medium$Pn.median))))
abline(0, 1)

plot(x = dat.mut$Pn.median, 
     y = dat.medium$Pn.median, 
     xlab = "mut.Pn.median", 
     ylab = "medium.Pn.median", 
     xlim = c(min(c(dat.mut$Pn.median, dat.medium$Pn.median)), max(c(dat.mut$Pn.median, dat.medium$Pn.median))),
     ylim = c(min(c(dat.mut$Pn.median, dat.medium$Pn.median)), max(c(dat.mut$Pn.median, dat.medium$Pn.median))))
abline(0, 1)
```

>>>>

---
title: "Structural Divergence - Protein dynamics"
author: "Maria Laura Marcos, Julian Echave"
output: html_document
---

# Introduction
In this report we compare Pn profiles with the fluctuation along normal modes.

@@

```{r set_file_names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))
site.info.fname = file.path(data.dir, paste(p.ref, "_m.da.ca.csv", sep = ""))

@@

### energy
energy.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.va.csv", sep = ""))
energy.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.va.csv", sep = ""))
energy.mut.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.va.csv", sep = ""))
```

@@

Pn.medium = read.csv(Pn.medium.fname, header = T)
rownames(Pn.medium) = paste("medium", rownames(Pn.medium), sep = "")
energy.medium = read.csv(energy.medium.fname, header = T)
rownames(energy.medium) = paste("medium", rownames(energy.medium), sep = "")
```

@@

modeComparison.exp = merge(energy.exp.long,Pn.exp.long,by=c("mode", "protein"))
modeComparison.exp = ddply(modeComparison.exp, "protein", mutate, "Pn" = Pn / mean(Pn, na.rm = T))

modeComparison.exp <- na.omit(modeComparison.exp)
modeComparison.medium <- na.omit(modeComparison.medium)
modeComparison.mut <- na.omit(modeComparison.mut)

# put simulated and experimental data together
modeComparison.all.datasets = rbind(data.frame("dataset" = "exp", modeComparison.exp),
                                    data.frame("dataset" = "mut", modeComparison.mut),
                                    data.frame("dataset" = "medium", modeComparison.medium))

modeComparison.all.datasets$dataset = factor(modeComparison.all.datasets$dataset, levels = c("exp", "mut", "medium"))
```

@@

dat = ddply(dat,c("dataset", "mode"), function(x) {
    energy.mean = mean(x$energy)
    Pn.mean = mean(x$Pn)
    ndata = sum(!is.na(x$Pn))
    Pn.se = sd(x$Pn, na.rm = T) / sqrt(ndata)
    data.frame(energy.mean, Pn.mean, Pn.se)
})

p = ggplot(dat, aes(x = mode, y = Pn.mean, col = dataset, fill = dataset)) +
    geom_line() +
    geom_ribbon(aes(ymin = Pn.mean - f95 * Pn.se, ymax = Pn.mean + Pn.se), alpha = .4, show.legend = F) +
    labs(x = "mode", y = "<Pn>") +
    scale_colour_manual(name = NULL, values = myColors) + 
    scale_fill_manual(name = NULL, values = myColors)

p = ggplot(dat, aes(x = (1/energy.mean), y = Pn.mean, col = dataset, fill = dataset)) +
    geom_point() +
    labs(x = "1/lambda_n", y = "<Pn>") +
    scale_colour_manual(name = NULL, values = myColors) + 
    scale_fill_manual(name = NULL, values = myColors)

@@

```{r correlation}
dat = subset(dat, mode < 101)
dat.exp = subset(dat, dataset == "exp")
dat.mut = subset(dat,dataset == "no-selection")
dat.medium = subset(dat,dataset == "selection")

r.exp = cor(dat.exp$Pn.mean, (1/dat.exp$energy.mean))
r.mut = cor(dat.mut$Pn.mean, (1/dat.mut$energy.mean))
r.medium = cor(dat.medium$Pn.mean, (1/dat.medium$energy.mean))
cor.df = data.frame("comparison" = c("exp", "mut", "medium"),
                    "R" = c(r.exp, r.mut, r.medium))
knitr::kable(cor.df, digits = 2)

# plots
par(pty = "s")

plot(y = dat.exp$Pn.mean, x = (1/dat.exp$energy.mean), xlab = "1/lambda_n", ylab = "<Pn> experimental")
fluc = (1/dat.exp$energy.mean)
lm = lm(dat.exp$Pn.mean ~ fluc)
abline(lm)
legend("topleft", paste ("r = ", round(r.exp, digits = 2)), bty = "n")

plot(y = dat.mut$Pn.mean, x = (1/dat.mut$energy.mean), xlab = "1/lambda_n", ylab = "<Pn> no-selection")
fluc = (1/dat.mut$energy.mean)
lm = lm(dat.mut$Pn.mean ~ fluc)
abline(lm)
legend("topleft", paste ("r = ", round(r.mut, digits = 2)), bty = "n")

plot(y = dat.medium$Pn.mean, x = (1/dat.medium$energy.mean), xlab = "1/lambda_n", ylab = "<Pn> selection")
fluc = (1/dat.medium$energy.mean)
lm = lm(dat.medium$Pn.mean ~ fluc)
abline(lm)
legend("topleft", paste ("r = ", round(r.medium, digits = 2)), bty = "n")
```
